---
title: "Rain Movies"
author: "Matthew Ross & John Kemper"
date: "11/15/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Using R to make movies

While most of the visualizations made in R are `static`, with a single image 
displaying the data, one can also make videos in R. These are particularly 
useful when data has a temporal component or a third axis for display. Here
we will show a few ways to make videos using R and airport weather data 

```{r setup, include=FALSE}
library(tidyverse)
library(riem) # Precipitation data
library(animation) #movie making
library(USAboundaries) # boundary data
library(gganimate) #another way to make movies
library(lubridate)
library(ggthemes)
library(tmap)
library(sf)
library(USAboundariesData)
library(RColorBrewer)
library(tmaptools)

theme_set(theme_few()) # Set ggplot theme


knitr::opts_chunk$set(echo = TRUE)
```


# Data acquisition and reshaping


## Finding our station names

```{r}
#riem_networks() # look for network names

co_stations <- riem_stations(network = 'CO_ASOS')

#View(co_stations)

stations <- c('AFF', 'FNL')

```

## Download station data

```{r}

af_p <- riem_measures(station='AFF',date_start = '2018-10-01',
                      date_end = '2019-09-30')


fc <- riem_measures(station='FNL',date_start = '2018-10-01',
                      date_end = '2019-09-30')

View(fc)



```


## Convert both to daily

```{r}

daily_p_fr <- fc %>% 
  bind_rows(af_p) %>%
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()


```


# Making videos 

## Boring first video

```{r}

slice_p <- daily_p %>%
  slice(1:20)

slice_p

p1 <- ggplot(slice_p,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_point() +
  labs(title = 'Date: {frame_time}', x='Date',y= 'Mean Temp (f)') + 
  transition_time(date) 


animate(p1, nframes=nrow(slice_p))


```


## Leaving a trace

```{r}
p2 <- ggplot(daily_p,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_line() + 
  geom_point() +
  labs(title = '100 mile temp diff', x='Date',y= 'Mean Temp (f)') + 
  transition_reveal(date) + 
  scale_color_few()


animate(p2, nframes=nrow(daily_p))



```


## Static background plot


```{r}
usa <- us_states()

co <- usa %>%
  filter(name == 'Colorado')

sf_p <- daily_p %>%
  st_as_sf(coords=c('lon','lat'),crs=4326)


co_elev <- elevatr::get_elev_raster(co,z=8)

tm_shape(co) + 
  tm_polygons() + 
  tm_shape(co_elev) +
  tm_raster(style='cont',palette = 'Greens') + 
  tm_shape(sf_p) + 
  tm_dots(col='station',size=3,palette=c('red','white'))

```



## Quick plot

```{r}

basemap <- tm_shape(co) + 
    tm_polygons() + 
  tm_shape(co_elev) +
    tm_raster(style='cont',
              palette = 'Greens',
              title='Elevation (m)') 

basemap
```



```{r,eval=F}
 
full_dates <- unique(sf_p$date)


sequence <- 1:10
saveVideo({
  for(i in sequence){
  sub_sf <- sf_p %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(col='station',
            size=2,
            palette = c('red','white'),
            legend.show = F) 
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i]) + 
          tm_add_legend(type='symbol',
                        labels=c('Air Force',
                                 'Fort Collins'),
                        col=c('red','white'),
                        title='Station'))
  }
},movie.name = 'precip_flash.mp4',
ffmpeg = "C:/Program Files/ffmpegx/bin/ffmpeg.exe",
  interval=0.05,
  ani.width=1000,
  ani.height=800)


```

# A cooler video

## Let's see if we can see weather fronts by downloading all Colorado data and
trying to get those to flash


```{r,eval=F}

all_co <- map(co_stations$id,~riem_measures(.x,date_start='2018-10-01',
                                         date_end = '2019-09-30'))
stack_co <- do.call('rbind',all_co)

daily_p_z <- stack_co %>% 
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()

save(daily_p,file='DailyP.RData')
```

## Full state video

This code is almost identical to the above, but it uses all ASOS/AWOS precip
stations


```{r, eval=F}
load('DailyP.RData')

daily_sf <- daily_p %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326,remove=F)




full_dates <- unique(daily_sf$date)


sequence <- 1:length(full_dates)
saveVideo({
  for(i in sequence){
  sub_sf <- daily_sf %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(title.size='Daily Precip (in)',
            col='lon',
            size='daily_p',
            style='cont',
            palette='Oranges',
            legend.show = F,
            size.max=19,
            scale=3.5,
            sizes.legend=c(0,1,5,10,20))
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i],
        legend.position=c(0.7,0.2)))
  }
},movie.name = 'precip_flash_all.mp4',
  interval=0.12,
  ani.width=1000,
  ani.height=800)



```



# In Class work.

## GGanimate

Make a video showing the correlation between Air Force Academy and Fort Collins
between the min, mean, and max temps. The movie should cover the full year. 

```{r}
#Here, we make an animated plot of max min and mean temperature at the Air Force Academy and Fort Collins

###Make a long dataframe where max, min, mean temperature are now a variable (named type)
daily_p_fr_long <- daily_p_fr %>%
  gather(., key = "type", value = "temperature", mean_temp:max_temp)


###function to remove "_" from facet wrap labels
remover <- function(string){
  
  str_replace_all(string, "_", " ")
  
}


###Construct plot of max, mean, min temp at 
###Air Force Academy and Fort Collins
maxmeanmin_plot <- daily_p_fr_long %>%
  ggplot() +
  geom_line(aes(x = date, y = temperature, color = station, linetype = type), size = 0.5) +
  theme_few() +
  scale_color_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c("solid", "longdash", "dotted")) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Temps at Air Force Academy and Fort Collins") +
  facet_wrap(~type, labeller = as_labeller(remover)) +
  guides(linetype = FALSE) +
  transition_reveal(date)


  
###Animate the plot
animate(maxmeanmin_plot, nframes=100)


```

## Make a video of mean temperature differences for the full Colorado dataset
(you need to download this [here](https://www.dropbox.com/s/s7zt18elgoh8q2b/DailyP.RData?dl=0)) first.

The video should look similar to the one above, but I want you to choose your own 
style (change the background, add other layers (like counties or cities)), etc...

```{r}

#load data
load('DailyP.RData')


#Here, we set up and make the basemap
###state outline
co <- us_states(states = "Colorado") 

###Colorado counties
co_counties <- us_counties(states = "Colorado") 

###Some Colorado cities
co_cities <- us_cities() %>%
  filter(state_name == "Colorado") %>%
  filter(city %in% c("Colorado Springs", "Denver", "Fort Collins", "Grand Junction", 
                   "Durango", "Pueblo", "Glenwood Springs"))

###Make hillshade
county_elev <- crop(co_elev, co_counties)
county_elev_slope <- terrain(county_elev, opt = "slope")
county_elev_aspect <- terrain(county_elev, opt = "aspect")
county_hillshade <- hillShade(county_elev_slope, county_elev_aspect, angle = 45, direction = 0)

  
###basemap
basemap_new <- tm_shape(county_hillshade) + 
  tm_raster(style = "cont", palette = "Greys", legend.show = FALSE) +
  tm_shape(co_counties) +
  tm_polygons(alpha = 0.5, border.alpha = 1, border.col = "white") +
  tm_shape(co_cities) +
  tm_symbols(size = 0.5, col = "black", shape = 15) +
  tm_text("city", xmod = -1, ymod = -0.7, fontfamily = "sans", size = 1.3) 

basemap_new

#Here, we make a movie my dudes
###Make daily data spatial
spatial_daily_p <- daily_p %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326,remove=F)


###Extract days from data
full_dates <- unique(spatial_daily_p$date)
full_dates

###Make video
sequence <- 1:length(full_dates)

saveVideo({
  
  for(i in sequence){
    
    
    sub_temp_sf <- spatial_daily_p %>%
      filter(date == full_dates[i])
    
    if(nrow(sub_temp_sf)>0) {
      
      finalmap_temperature <- basemap_new +
      tm_shape(sub_temp_sf) +
      tm_symbols(col = "mean_temp", 
                 palette = viridis::plasma(n = 50),
                 title.col = " Mean Temperature (F)")
    }
    
    
    else{finalmap_temperature <- basemap_new}
    
    print(finalmap_temperature + 
            tm_layout(title = as.character(full_dates[i]),
                      title.size = 2.5,
                      legend.text.size = 1.3,
                      legend.text.fontfamily = "sans",
                      legend.title.size = 1.5,
                      legend.title.fontfamily = "sans"))
    
  }
    
  
},movie.name = "mean_temp.mp4",
  interval=0.12,
  ani.width=1000,
  ani.height=800,
  ffmpeg = "C:/Program Files/ffmpegx/bin/ffmpeg.exe")

```

